Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-08-30 Master Thesis/introduction.tex
Rewrite of abstract and introduction

Finite fields have an abelian absolute Galois group. In fact, they all have the same absolute Galois group $\widehat{\mathbb{Z}}$, therefore not much information can be retrieved from these groups. The study of these types of questions belong to the area of \textit{anabelian geometry}. This term refers to the fact that the less abelian $\pi_1^\text{ét}(X)$ is, the more information it carries about $X$.

Number fields stand in stark contrast to finite fields. Two number fields $k$, $k'$ with isomorphic Galois groups $G_k$, $G_{k'}$ are isomorphic as fields as well. It was J. Neukirch who investigated this question for the first time in the end of the 1960s (see \cite{Ne69}, \cite{Uc76}, \cite{Ne77}, \cite{Uc77}). Together with K. Uchida, their studies resulted in the following theorem for global fields:

@@

It is believed that spectra of finitely generated fields are the building blocks of the anabelian world, which led Grothendieck to the conjecture that Theorem \ref{thm:neukirch-uchida} should also hold for arbitrary finitely generated fields $k$ and $k'$ of characteristic $0$.

This has been proven by F. Pop (see \cite{Po90}, \cite{Po94}, \cite{Po95}) using model theory. In this thesis, we want to follow M. Spiess \cite{Sp96} to prove the statement for finitely generated fields with transcendence degree $1$ over $\mathbb{Q}$, using a purely arithmetical approach and a direct generalization of the original proof of Theorem \ref{thm:neukirch-uchida}:

@@

I also wish to offer my thanks to all my friends, who supported me in this endeavor and encouraged me to go on. Last but not least, I am grateful for my parents who never ran out of patience with me.

This thesis was hard, arduous work, but, in the end, also very satisfying. And I couldn't have done it without you all. Thank you.

>>>>

Finite fields have an abelian absolute Galois group. In fact, they all have the same absolute Galois group $\widehat{\mathbb{Z}}$, therefore not much information can be retrieved from these groups. The study of these type of questions belong to the area of \textit{anabelian geometry}. This term refers to the fact that the less abelian $\pi_1^\text{ét}(X)$ is, the more information it carries about $X$.

Number fields stand in stark contrast to finite fields. Two number fields $k$, $k'$ with isomorphic absolute Galois groups $G_k$, $G_{k'}$ are isomorphic as fields as well. This result is due to J. Neukirch who investigated these type of questions for the first time in the end of the 1960s (see \cite{Ne69}, \cite{Ne77}). With K. Uchida's studies (see \cite{Uc77}), it resulted in the following theorem for global fields:

@@

It is believed that spectra of finitely generated fields are the building blocks of the anabelian world, which led to the conjecture that Theorem \ref{thm:neukirch-uchida} should also hold for arbitrary infinite fields $k$ and $k'$ which are finitely generated over their prime fields. This has been proven by F. Pop using model theory (see \cite{Po90}, \cite{Po94}, \cite{Po95}).

In this thesis, we want to follow M. Spieß \cite{Sp96} to prove the statement for finitely generated fields with transcendence degree $1$ over $\mathbb{Q}$, using a purely arithmetical approach, which is a direct generalization of the original proof of Theorem \ref{thm:neukirch-uchida}:

@@

I also wish to offer my sincerest thanks to all my friends, who supported me in this endeavor and encouraged me to go on. Last but not least, I am grateful for my parents who never ran out of patience with me.

This thesis was hard, arduous work, but, in the end, also very satisfying. And I couldn't have done it without all of you. Thank you.

### 2017-08-30 Master Thesis/synopsis.tex
Rewrite of abstract and introduction

\chapter{Correspondences}

Let $K,K'$ be two finitely generated fields of characteristic $0$ and dimension $2$ and let $\sigma:G_K\to G_{K'}$ be an isomorphism of profinite groups. For intermediate fields $L$ of $\overline{K}/K$ we denote the corresponding intermediate field of $\overline{K'}/K'$ by $L'$. In other words, $\sigma (G_L)=G_{L'}$.

>>>>

\chapter{Partial Correspondences}

Let $K,K'$ be two finitely generated fields of characteristic $0$ and dimension $2$, and let $\sigma:G_K\to G_{K'}$ be an isomorphism of profinite groups. For intermediate fields $L$ of $\overline{K}/K$ we denote the corresponding intermediate field of $\overline{K'}/K'$ by $L'$. In other words, $\sigma (G_L)=G_{L'}$.

### 2017-08-30 Master Thesis/thesis.tex
Slightly increase linespread
Rewrite of abstract and introduction

\documentclass[11pt, a4paper, openany]{memoir}
\usepackage{fontspec}
\usepackage[english]{babel}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{paralist, tikz}
\usetikzlibrary{cd, babel}
\usepackage[inner=4cm, outer=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{hyperref}
\hypersetup{bookmarksopen=true, pdfstartview=FitH}
\linespread{1.1}
\author{Yichuan Shen}
\title{Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields}

@@

\vspace*{0cm}
\vfill
\renewcommand{\abstractname}{Abstract}
\begin{abstract}
In this thesis we will follow M. Spiess' argument and present his purely arithmetic proof of the fact that isomorphisms between the absolute Galois groups of function fields of transcendence degree $1$ over number fields are induced by field isomorphisms between these fields.
\end{abstract}

@@

\renewcommand{\abstractname}{Zusammenfassung}
\begin{abstract}
In dieser Arbeit wollen wir M. Spiess' rein arithmetischer Beweis nachgehen und zeigen, wieso Isomorphismen zwischen den absoluten Galoisgruppen von Funktionenkörpern vom Transzendenzgrad $1$ über Zahlkörper schon von Körperisomorphismen zwischen diesen Körpern induziert werden.
\end{abstract}
\vfill
\clearpage

>>>>

\documentclass[11pt, a4paper, openany]{memoir}
\usepackage{fontspec}
\usepackage[english]{babel}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{paralist, tikz}
\usetikzlibrary{cd, babel}
\usepackage[inner=4cm, outer=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{hyperref}
\hypersetup{bookmarksopen=true, pdfstartview=FitH}
\linespread{1.15}
\author{Yichuan Shen}
\title{Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields}

@@

\vspace*{0cm}
\vfill
\renewcommand{\abstractname}{Abstract}
\begin{abstract}
In this thesis we will follow M. Spieß' argument and present his purely arithmetic proof of the fact that isomorphisms between the absolute Galois groups of function fields of transcendence degree $1$ over number fields are induced by field isomorphisms between these fields.
\end{abstract}

@@

\renewcommand{\abstractname}{Zusammenfassung}
\begin{abstract}
In dieser Arbeit wollen wir zeigen, wieso Isomorphismen zwischen den absoluten Galoisgruppen von Funktionenkörpern vom Transzendenzgrad $1$ über Zahlkörper schon von Körperisomorphismen zwischen diesen Körpern induziert werden, indem wir M. Spieß' rein arithmetischen Beweis nachgehen.
\end{abstract}
\vfill
\clearpage
